Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-07-04 GNproof.tex
attempt to write the GN proof

\begin{document}
\maketitle

\section{Introduction}

This paper is the second in a two part series investigating the combinatorics and geometry underlying SYM $N=4$ theory. This series lays out several results about the relationship between the Wilson loop diagrams and the positroid cells that give a CW complex structure to the positive Grassmanian. This paper is concerned with identifying which positroid cells correspond to which Wilson loop diagrams, and how this translates to a geometric tiling of a subspace of the positive Grassmannians.

% Wilson loop diagrams associate a volume form to a positroid cell.

% In this paper, we give an algorithm to identify which cell corresponds to which diagram, via Grassmann Necklaces. Using the Le diagram representation of positroid cells, we prove results about the dimensions of these cells. Finally, we state some results about the structures of the volume forms defined on these positroid cells by the Wilson loop diagrams.

In contrast, the first paper in this series focuses on combinatorial and matroidal properties of the Wilson loop diagrams. The main focus of that paper was to enumerate the positroid cells that could be associated to Wilson loop diagrams with a fixed number of vertices and propagators, as well as counting the number of Wilson loop diagrams in the pre-image of any positroid cell under this association.

In recent years, there has been significant interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, galashinlam, AmplituhedronDecomposition}. This started with the observation that BCFW diagrams, which represent the on shell interactions of this theory, correspond to plabic graphs \cite{GrassmannAmplitudebook}. Plabic graphs were introduced by Postnikov in \cite{Postnikov}, and provide one of the many equivalent ways to classify the positroid cells of the positive real Grassmannian $\Gr(k,n)$. This led to the representation of the on shell integrals associated to this theory in terms of a geometric space (the Amplituhedron) embedded in $\Gr(k,n)$ \cite{Arkani-Hamed:2013kca,Arkani-Hamed:2013jha}. Moreover, integrals associated to the BCFW diagrams define a volume form on the Amplituhedron, \hlfix{which is useful for ...}{we need a vague one-line description here}. Since then, there has been significant work to understand the volume form and its physical implications \cite{UnwindingAmplituhedron,galashinlam}.

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams, and adding up the corresponding sums of Feynman integrals associates a volume form to \hlfix{space}{I don't know if this was supposed to say ``space'' or ``this space'', and if ``this space'' I don't know which space it's referring to.} \cite{Hodges:2012ym}. The first author and Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$ \cite{wilsonloop}, while other work has shown a connection between the space parametrized by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}.

In the above work and elsewhere, it is clear that while the space parametrized by the Wilson loop diagrams is related to the Amplituhedron, they are not the same space. For instance, it is conjectured that the Amplituhedron is orientable, whereas the first author and Marcott show in \cite{non-orientable} that the space parametrized by the Wilson loop diagrams can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, as one sees from counting arguments in comparing the results of \cite{AmplituhedronDecomposition} and \cite{casestudy}. This is not too alarming, as \hlfix{one expects}{can we call this a conjecture?} the space parametrized by the Wilson loop diagrams to project down to something close to the Amplituhedron \cite{wilsonloop, Amplituhedronsquared}.

This paper continues this work by identifying which positroid cells are associated to which Wilson loop diagram: specifically, we give an algorithm to read off the Grassmann necklace of the associated positroid cell from the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$ dimensional; by contrast, the BCFW cells are all $4k$ dimensional \hlfix{[ref]}{someone actually proved this, right?}. Again, this is not too alarming, as it was shown in \cite{non-orientable} that the actual geometric space parametrized by the Wilson loop diagrams is a non-orientable $k$-dimensional bundle over this $3k$ dimensional manifold in $\Gr(k,n)$.

As noted above, the sum of the Feynman integrals associated to Wilson loop diagrams gives the volume of \hlfix{the space}{help?}; if this volume is to be finite, then any singularities appearing in individual integrals must cancel out in the overall sum. This has been verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general due to the fact that the relationship between Wilson loop diagrams and $3k$ dimensional positroid cells of $\Gr(k,n)$ is neither one-to-one nor onto. In this paper, we give a characterization of the factors appearing in the denominator of these Feynman integrals in terms of the Grassmann necklace of the associated positroid, thus providing a much simpler and more tractable link between the singularities and the diagrams.

% when computing a finite volume to a geometric object defined by the Wilson loop diagrams, one must ensure that any singularities cancel. This has been verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it has been hard to prove in general due to the fact that the relationship between Wilson loop diagrams and $3k$ dimensional positroid cells of $\Gr(k,n)$ is neither one-to-one nor onto. In this paper, we give a characterisation of the denominator of the Feynman integral defined by a Wilson loop diagram in terms of the algorithm that associated the same diagram to a Grassman Necklace.

\subsection{Roadmap}

In Section~\ref{section general background}, we summarize the required background for Wilson loop diagrams (subsection~\ref{section WLD background}), the theory of positroids in $\Gr(k,n)$ and their characterization in terms of Grassmann necklaces and Le diagrams (subsection~\ref{sec:positroid background}), and the link between Wilson loop diagrams and positroids (subsection~\ref{sec:WLD as positroids}).

In Section~\ref{sec GN algorithm} we state an algorithm for constructing the Grassmann necklace of the positroid associated to a Wilson loop diagram (Algorithm~\ref{alg:put GN on WLD}) and prove its correctness (Theorem~\ref{res:alg gives GN}). While it was already possible to construct the Grassmann necklace of a Wilson loop diagram via existing bijections between some of the various combinatorial objects that index the positroid cells, this process was convoluted and involved multiple steps; our algorithm allows us to track directly how each propagator in the Wilson loop diagram contributes to each term in the Grassmann necklace. Many of the technical lemmas required to prove Theorem~\ref{res:alg gives GN} continue to be important in subsequent sections: in particular we have Lemma~\ref{lem sian}, which identifies certain configurations of propagators which must appear in any admissible Wilson loop diagram, and Lemma~\ref{vertex cyclic int lem}, which shows that the pattern of values that a propagator contributes to each term of the Grassmann necklace is both simple and predictable.

In Section~\ref{sec dim} we examine the dimension of positroids associated to Wilson loop diagrams, and show that the dimension is always equal to three times the number of propagators in the diagram (Theorem~\ref{thm dim}).  While this has been proved by Marcott in \hlfix{[ref]}{Did this get published yet?} using geometric methods, our proof is constructive and explicitly relates the position of the plusses in the Le diagram to the propagators in the Wilson loop diagram.

Finally in Section~\ref{sec poles} we characterize the denominator of the integral of a Wilson loop diagram in terms of how propagators contribute to the Grassmann necklace.  This characterization is given explicitly by Algorithm~\ref{alg WLD to denom via GN} and Theorem~\ref{thm denom}.

\section{Background}\label{section general background}

\subsection{Wilson Loop diagrams}\label{section WLD background}

\begin{dfn}\label{WLdfn}
A {\bf Wilson loop diagram} $W = (\cP,V)$ consists of a cyclically ordered set $V$ with one distinguished element (this plays the role of the ``first'' element when we need a compatible linear ordering on $V$), and a set $\cP \subseteq V \times V$. The elements of $V$ are the {\bf vertices} of $W$, and elements of $\cP$ are the {\bf propagators} of $W$. \end{dfn}

A Wilson loop diagram is depicted as a circle, with the vertices $V$ arranged along the edge of the circle and listed counterclockwise in their cyclic order. The arc between two consecutive vertices is referred to as an {\bf edge} of the diagram, and the $i$th edge of $W$ is the edge from vertex $i$ to its successor. Each propagator $(i,j) \in \cP$ is denoted by a wavy line inside the circle joining edge $i$ and edge $j$.

\begin{figure}[h]
\[W\ =\ \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
    \drawprop{5}{0}{8}{0}
    \drawprop{5}{1}{7}{0}
		\end{scope}
	\end{tikzpicture}\]

\caption{The Wilson loop diagram $W = \big(\{(1,4), (2,4), (5,7), (5,8)\},[8]\big)$ which has four propagators and eight vertices.}
\label{fig:ex of WLD}
\end{figure}

Let $[n]$ denote the set of integers $\{1,\dots,n\}$ with the obvious cyclic and linear orders. Our primary focus is Wilson loop diagrams with $V = [n]$, so in order to simplify the exposition we will often use $i+1$ as shorthand for ``the successor of vertex $i$''.

Generally speaking the propagators in a Wilson loop diagram are undirected, so $(i,j)$ and $(j,i)$ represent the same propagator. If we need to impose a direction on a propagator $p = (i,j)$, for example in order to consider the region ``inside'' or ``outside'' of $p$, we write $(p,<_i)$ to indicate that $p$ is directed from edge $i$ to edge $j$, and $(p,<_j)$ for the opposite direction.

We also impose the convention that all vertices should always be interpreted in terms of the cyclic order on $V$; for example, if $V = [n]$ then $(n,n+3)$ means the propagator connecting edges $3$ and $n$. This viewpoint is beneficial for us since many of our proofs rely on restricting our attention to a region of the diagram bounded by a given propagator, or to short propagators such as $(i,i+2)$, and these behave the same regardless of whether we cross the ``first'' vertex or not; however, we emphasize that this is different to the standard convention in the physics literature, where propagators are written $(i,j)$ with $1 \leq i<j \leq n$.

\begin{dfn} \label{VPropdfn}
Let $W = (\cP, V)$ be a Wilson loop diagram.
\begin{enumerate}
\item For $p = (i,j) \in \cP$, let $V(p) = \{i, i+1, j, j+1\}$ denote the set of {\bf vertices supporting $p$}. For a set of propagators $P \subseteq \cP$, define 
\[V(P) = \bigcup_{p \in P} V(p)\]
to be the vertex support of $P$.
\item For $U \subseteq V$, the set of {\bf propagators supported on $U$} is denoted by
\[\Prop(U) = \{ p \in \cP\ |\ V(p) \cap U \neq \emptyset \}.\]
Note that a propagator in $\Prop(U)$ does not need to have its entire support contained in $U$. 
\item Vertices of $W$ that do not support any propagators are called {\bf non-supporting}. The set of all non-supporting vertices in $W$ is denoted $F(\emptyset)$. 
\end{enumerate}
\end{dfn}

In particular, we will see in later sections that all four vertices in the support of a propagator are important in understanding the contribution of that propagator to the overall diagram. This motivates the following definition of a subdiagram, which can consist of any subset $Q$ of the propagators and any subset $U$ of the vertices of $W$ {\em as long as we retain all vertices in the support of $Q$}.

\begin{dfn} \label{subdiagramdfn}
Let $W = (\cP, V)$ be a Wilson loop diagram. Then $W' = (Q,U)$ is a {\bf subdiagram} of $W$ if $Q \subseteq \cP$ and $V(Q) \subseteq U \subseteq V$.
\end{dfn}

With these definitions in hand, we can impose some conditions on the density and behavior of the propagators of Wilson loop diagrams.

\begin{dfn}\label{admisdfn}
A Wilson loop diagram $W = (\cP, V)$ is called {\bf admissible} if:
\begin{enumerate}
\item $|V| \geq |\cP| + 4$.
\item There does not exist a non-empty set of propagators $Q \subseteq \cP$ such that $|V(Q)| < |Q| + 3$.
\item There does not exist a pair of propagators $\{(i,j),(k,l)\} \subseteq \cP$ such that $i < k < j < l$.
\end{enumerate}
 \end{dfn}

The first condition bounds the total number of propagators in the diagram, while the second limits how densely propagators can be fitted into any part of the diagram; in particular, it prohibits propagators that connect two adjacent edges and parallel propagators that start and end on the same pair of edges. The third condition simply requires that propagators are non-crossing.

These conditions are imposed by the physical interpretation of Wilson loop diagrams (see for example \cite{Adamo:2011pv,Adamo:2012xe,wilsonloop,LipsteinMason}). We therefore restrict our attention in this paper to admissible Wilson loop diagrams and their subdiagrams only.

Note that a subdiagram $(Q,U)$ of an admissible Wilson loop diagram need not be admissible itself: it inherits conditions 2 and 3 automatically, but we could have $|U| = |Q| + 3$. If a Wilson loop diagrams satisfies conditions 2 and 3 of Definition~\ref{admisdfn}, we call it {\bf weakly admissible}.

These diagrams correspond to (tree level) amplitudes of SYM $N=4$ theory in twistor space. As such, it associates a real value to $n$ particles, one for each vertex. Each particle is represented as a section of a $|\cP|$-vector bundle over twistor space, projected onto a real subspace in a process called bosonization \cite{Arkani-Hamed:2013jha}. In other words, each Wilson loop diagram defines a functional on external particles. This is represented as an integral $I(W)$, which is the equivalent of a Feynman integral in this setting.

A $|\cP| \times n$ matrix $C(W)$ represents the (tree level) particle interactions. Each row of $C(W)$ corresponds to an internal propagator, each column to an external particle. $C(W)$ is defined as follows:

\begin{equation} C(W)_{p,q} = \begin{cases} c_{p,q} & \textrm{ if } q \in V(p) \\
0  & \textrm{ if } q \not \in V(p)  \end{cases}
\;. \label{C(W) dfn}\end{equation} \todo{$c_{p,q}$ or $x_{p,q}$? Should be consistent}

The entries $c_{p,q}$ are non-zero real variables. Thus the zero entries in row $p$ correspond to the external particles that the propagator $p$ does not interact with. See \ref{integraldetails} for an example of a diagram $W$ and its associated matrix.

The matrix $C(W)$ can be interpreted as a parametrization of the positroid cell in the positive Grassmannian $\Gr(n, |\cP|)$ associated to the $W$, which we denote by $\Sigma(W)$ \cite{wilsonloop}. Most of the work in this paper, and in \cite{generalcombinatoricsI} focuses on characterizing the positroid cell assoicated to eash Wilson loop diagram $W$.

Note that since there is no ordering on the propagators of $W$, $C(W)$ is only defined up to rearrangement of the rows. This will not matter here, as it does not affect the parametrization of $\Sigma(W)$.

One may also interpret the functional $I(W)$ as associating a volume to each cell $\Sigma(W)$ \cite{wilsonloop, Amplituhedronsquared, HeslopStewart}. The volume form is written in terms of the elements of $C(W)$ as  \[ \Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V_{p_r}} \textrm{d}c_{p_r}}{R(W)} \;. \]
The interested reader is referred to \cite{Adamo:2012xe,HeslopStewart,LipsteinMason} for more information about the volume form.

For a fixed number of propagators and particles, the tree level amplitude of the Wilson loop diagrams is computed as the sum of the $I(W)$ for all $W$ of the correct size. By \cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha} this sum is finite for all sets of external particles, i.e. we have \bas \mathcal{A}_{k, n}^{tree} = \sum_{(\cP, [n]), |\cP| = k} I(W) <\infty. \eas

In particular, this implies that any singularities in the integrals $I(W)$ must cancel out in the sum. In particular, they must cancel on the boundaries of the positroid cells $\Sigma(W)$, analogous to the analysis in \cite{Arkani-Hamed:2013jha}. This has been checked for particular classes of examples (e.g. \cite{casestudy, Amplituhedronsquared, HeslopStewart}) but has not been proven in general.

With this motivation in mind, in this paper we study the denominator $R(W)$ of the volume form $\Omega(W)$, which we define next.

\begin{dfn}\label{def R(W)}
Let $W = (\cP,V)$ be an admissible Wilson loop diagram, and fix an edge $e$. Let $\{p_1 \ldots p_r\}$ be the propagators with one end lying on edge $e$, ordered so that $p_1$ is closest to the vertex $e$, $p_r$ closest to $e+1$, and $p_i$ is closer to $e$ than $p_{i+1}$ is. Define 
\[ R_e =  c_{p_1,e+1} \prod_{j= 1}^{r-1} \left((c_{p_j,e} c_{p_{j+1},e+1} - c_{p_{j+1},e} c_{p_{j},e+1} ) \right) c_{p_r,e}\;\]
to be the component of $R(W)$ associated to edge $e$. Note that if $r = 1$ (i.e. there is exactly one propagator lying on edge $e$) then this expression simplifies to $R_e = c_{p,e} c_{p,e+1}$. If $r=0$, set $R_e = 1$. The denominator $R(W)$ is defined to be the product of all of the $R_e$, i.e. 
\[R(W) = \prod_{e \in V} R_e.\]
\end{dfn}

In Section \ref{sec poles}, we show that the denominator $R(W)$ is determined by the positroid cell $\Sigma(W)$ and the structure of $W$, as encapsulated by $C(W)$. In fact, it can be read directly off of $C(W)$, in terms of characterization of $\Sigma(W)$ called the Grassmann necklace (see Definition~\ref{def:grassmann necklace}). We also give an algorithm to read $R(W)$ directly from the Wilson loop diagram, paving the way for a general study of these poles in future.

We end this section with an example of computing $C(W)$ and $R(W)$ for a particular Wilson loop diagram $W$.

\begin{eg} \label{integraldetails}
Consider the admissible Wilson loop diagram 
\[W = \big(\{(1,4), (2,4), (5,7), (5,8)\},[8]\big),\]
from Figure~\ref{fig:ex of WLD}. Ordering the propagators as listed above, we obtain the associated matrix
\[ C(W) = \left(
\begin{array}{cccccccc}
c_{1,1} & c_{1,2} & 0 & c_{1,4} & c_{1,5} & 0 & 0 & 0 \\
0 & c_{2,2} & c_{2,3} & c_{2,4} & c_{2,5} & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & c_{3,5} & c_{3,6} & c_{3,7} & c_{3,8} \\
c_{4,1} & 0 & 0 & 0 & c_{4,5} & c_{4,6} & 0 & c_{4,8}  \\
\end{array}
\right) \;.\]

To calculate $R(W)$, write \bmls R_1 = c_{1,1} c_{1, 2}\quad R_2 = c_{2,2} c_{2,3} \quad R_4 = c_{2,5} (c_{2,4}c_{1,5} - c_{2,5}c_{1,4})c_{1,4} \\ R_5 = c_{4,6} (c_{4,6}c_{3,5} - c_{3,6}c_{4,5})c_{3,5} \quad R_7 = c_{3,7} c_{3,8} \quad R_8 = c_{4,8} c_{4,1} \;.\emls All other $R_e$ are $1$.  Finally, the integrand associated to this Wilson loop diagram is \bas \Omega(W) = \frac{\textrm{d}c_{1,1}\textrm{d}c_{1,1}\textrm{d}c_{1,1}\textrm{d}c_{1,2}\textrm{d}c_{1,4}\textrm{d}c_{1,5}\textrm{d}c_{2,1}\textrm{d}c_{2,3}\textrm{d}c_{2,4}\textrm{d}c_{2,5}\textrm{d}c_{3,5}\textrm{d}c_{3,6}\textrm{d}c_{3,7}\textrm{d}c_{3,7}\textrm{d}c_{4,5}\textrm{d}c_{4,6}\textrm{d}c_{4,8}\textrm{d}c_{4,1}}{c_{1,1} c_{1, 2}c_{2,2} c_{2,3}c_{2,5} (c_{2,4}c_{1,5} - c_{2,5}c_{1,4})c_{1,4}c_{4,6} (c_{4,6}c_{3,5} - c_{3,6}c_{4,5})c_{3,5}c_{3,7} c_{3,8}c_{4,8} c_{4,1}} \;. \eas
\note{Is the numerator just supposed to be the product of $dc_{pq}$ for each $c_{pq}$ in the matrix? Also I think $c_{46}c_{35} - c_{36}c_{45}$ in the denominator is the wrong way around?}
\end{eg}

\subsection{Positroids, Grassmann necklaces, and Le diagrams}\label{sec:positroid background}

We summarize here only the subset of positroid theory that we require in this paper; the interested reader is referred to \cite{Postnikov} for more details.

Let $\binom{[n]}{k}$ denote the set of all $k$-subsets of $[n]$.

For our purposes, a {\bf positroid} is a matroid $M$ (with ground set $[n]$ and bases $\cB \subseteq \binom{[n]}{k}$) which can be represented by an element of the positive Grassmannian $\Gr(k,n)$. In other words, there exists a full-rank $k\times n$ real matrix whose $k\times k$ minors are all nonnegative and such that the minor $[[k]|J]$ is positive if and only if $J \in \cB$.

Postnikov shows in \cite{Postnikov} that the positroids of $\Gr(k,n)$ are indexed by many different collections of objects, each with their own advantages and disadvantages. The two most suited to our purposes are Grassmann necklaces and Le diagrams, which we introduce below. In order to do so, we first need some preliminary definitions.

For each $j \in [n]$, we can define a total order $<_j$ on the interval $[n]$ by
\[ j <_j j+1 <_j \dots <_j n <_j 1 \dots <_j j-1\;.\]
This in turn induces a total order on $\binom{[n]}{k}$, namely the lexicographic order with respect to $<_j$.  It also induces a separate partial order $\gale{j}$ on $\binom{[n]}{k}$ (the {\bf Gale order} \cite{Gale}), which is defined as follows: for 
\[A = \{a_1 <_j a_2 <_j \dots <_j a_k\} \text{ and } B = \{b_1 <_j b_2 <_j \dots <_j b_k\} \in \binom{[n]}{k},\] we define
\[A \gale{j} B \text{ if and only if } a_r \leq_j b_r \text{ for all }1 \leq r \leq k.\]
For example, in $\binom{[6]}{3}$ we have $\{2,5,6\}\gale{2} \{2,6,1\}$ but $\{2,5,6\}\not\gale{2}\{3,4,6\}$.

\begin{dfn}\label{def:grassmann necklace}
A {\bf Grassmann necklace} of type $(k,n)$ is a sequence $(I_1, \dots, I_n)$ of $n$ sets $I_i \in \binom{[n]}{k}$ such that for each $i \in [n]$:
\begin{itemize}
\item if $i \in I_i$, then $I_{i+1} = \big(I_i \backslash \{i\}\big) \cup \{j\}$ for some $j \in[n]$.
\item if $i \not\in I_i$, then $I_{i+1} = I_i$.
\end{itemize}
By convention, we set $I_{n+1} = I_1$.
\end{dfn}

By \cite[Theorem 17.1]{Postnikov}, the Grassmann necklaces of type $(k,n)$ are in 1-1 correspondence with the positroid cells in $\Gr(k,n)$. Each term $I_i$ is simply the minimal (with respect to the $<_i$-lex order) basis of the positroid. A characterization of all bases of the positroid in terms of the Grassmann necklace and the Gale order was given by Oh in \cite[Theorem 8]{Oh}: if $(I_1, \dots, I_n)$ is the Grassmann necklace associated to a positroid $M = ([n],\cB)$, then the bases of $M$ are exactly
\ba \cB = \left\{J \in \binom{[n]}{k}\ |\ I_i \gale{i} J \ \forall i \in [n]\right\}. \label{basesofmatroids}\ea
Thus the Grassmann necklace is well suited to testing whether a given $k$-set is a basis for $M$, and for generating a list of all bases of $M$.

\begin{dfn}\label{def:le diagram}
A {\bf Le diagram} is a Young diagram in which every square contains either a $+$ or a $0$, subject to the rule that if a square contains a $0$ then either all squares to its left (in the same row) must also contain a $0$, or all squares above it (in the same column) mustbasesofmatroids\ also contain a $0$, or both.
\end{dfn}

By \cite[Theorem 6.5]{Postnikov}, the set of all Le diagrams that fit within a $k\times(n-k)$ rectangle is in 1-1 correspondence with the positroid cells of $\Gr(k,n)$. Le diagrams are particularly useful for comparing dimensions of positroids, since the dimension of a positroid is equal to the number of $+$ squares in its Le diagram \cite[Theorem 6.5]{Postnikov}.

The rows and columns of a Le diagram are labelled as follows: given a Le diagram fitting inside a $k\times (n-k)$ box, arrange the numbers $1,2, \dots, n$ along its southeast border, starting from the top-right corner. See Figure \ref{fig:row column numbering} for examples.

\begin{figure}[h!]
\[\begin{tikzpicture}[baseline=(current bounding box.east),scale=0.75]
\draw (1,0) grid (2,3);
\draw (2,1) grid (4,3);
\draw (4,2) grid (5,3);
\draw[dotted] (2,0) -- (5,0) -- (5,2);

\node at (5.15,2.5) {\footnotesize$1$};
\node at (4.5,1.85) {\footnotesize$2$};
\node at (4.15,1.5) {\footnotesize$3$};
\node at (3.5,0.85) {\footnotesize$4$};
\node at (2.5,0.85) {\footnotesize$5$};
\node at (2.15,0.5) {\footnotesize$6$};
\node at (1.5,-0.15) {\footnotesize$7$};
%\node at (0.5,-0.15) {\footnotesize$8$};
\end{tikzpicture}
\qquad \qquad
\begin{tikzpicture}[baseline=(current bounding box.east),scale=0.75]
\draw (1,0) grid (2,3);
\draw (2,1) grid (4,3);
\draw (4,2) grid (5,3);
\draw[dotted] (2,0) -- (6,0) -- (6,3) -- (5,3);

\node at (5.5,2.85) {\footnotesize$1$};
\node at (5.15,2.5) {\footnotesize$2$};
\node at (4.5,1.85) {\footnotesize$3$};
\node at (4.15,1.5) {\footnotesize$4$};
\node at (3.5,0.85) {\footnotesize$5$};
\node at (2.5,0.85) {\footnotesize$6$};
\node at (2.15,0.5) {\footnotesize$7$};
\node at (1.5,-0.15) {\footnotesize$8$};
%\node at (0.5,-0.15) {\footnotesize$9$};
\end{tikzpicture}
\]
\caption{Row and column numbering for a Young diagram with $k = 3$, $n = 7$ (left) and $k = 3$, $n = 8$ (right). The top left box has coordinates $(1,7)$ in the left diagram, and $(2,8)$ in the right diagram.}
\label{fig:row column numbering}
\end{figure}

Specifying a $k$-subset $J \subseteq [n]$ therefore uniquely determines the shape of the Le diagram, by taking the elements of $J$ to be the row indices of the diagram.

An algorithm for constructing the Le diagram associated to a Grassmann necklace was given by Agarwala and Fryer in \cite{reversingOh}. Since we will make use of this algorithm in Section~\ref{sec dim} below, we summarize the process here.
\begin{algorithm}\label{alg:GN to Le} \ \cite[Algorithm 2]{reversingOh}
Let $(I_1,\dots,I_n)$ be a Grassmann necklace of type $(k,n)$. Within a $k \times(n-k)$ square, draw the Young diagram whose rows are labelled by $I_1$ (as per the convention above).

For each $i$, $2 \leq i \leq n$:
\begin{itemize}
\item Write \[I_1 \setminus I_i = \{a_1 > a_2 > \dots > a_r\}, \quad I_i \setminus I_1 = \{b_1 < b_2 < \dots < b_r\},\]
where the inequalities denote the $<_1$ order (subscripts suppressed for clarity).
\item For $1 \leq j \leq r$ , place a $+$ in square $(a_j,b_j)$ of the diagram. (We will sometimes refer to this $+$ as being {\em in the $a_j \rightarrow b_j$ position}.)
\end{itemize}
After performing the above for $2 \leq i \leq n$, place a $0$ in any remaining unfilled boxes.
\end{algorithm}

An algorithm for constructing the Grassmann necklace of a Le diagram also exists; this was given by Oh in \cite{Oh}. A method for using the Le diagram to test whether a given $k$-subset is a basis for the corresponding positroid or not was given by Casteels in \cite{CasteelsPaths}.

\subsection{Wilson loop diagrams as positroids}\label{sec:WLD as positroids}

Let $W = (\cP,[n])$ be an admissible Wilson loop diagram with $k$ propagators, and $C(W)$ the associated matrix defined in \eqref{C(W) dfn}. Let $M(W)$ be the matroid realized by $C(W)$, i.e. the matroid with ground set $[n]$ whose independent sets are exactly the sets $V \subseteq [n]$ such that the columns of $C(W)$ indexed by $V$ are linearly independent.

In \cite{wilsonloop}, Agarwala and Amat show that $M(W)$ is also characterized by the Wilson loop diagram itself:

\begin{thm} \label{thm WLD defines matroid} \cite[Theorem 3.6]{wilsonloop} The independent sets of the matroid $M(W)$ associated to an admissible Wilson loop diagram $W = (\cP,[n])$ and realized by $C(W)$ are exactly those subsets $V \subseteq [n]$ such that $\nexists U\subseteq V$ satisfying $|\Prop(U)| < |U|$. \end{thm}
In other words, the independent sets of $M(W)$ correspond to sets of vertices in $W$ in which no subset supports fewer propagators than the vertices it contains.

One useful corollary of this, which we will want to keep in mind later, is the following.
\begin{cor}\label{lem basis as perm}
Let $W = (\cP,[n])$ be an admissible Wilson loop diagram and let $M(W)$ be its associated matroid. A subset $J \subseteq [n]$ is an independent set of $M(W)$ if and only if there exists an injective set map $f : J \rightarrow \cP$ with the property that for each $j\in J$ we have $j \in V(f(j))$.
\end{cor}

\begin{proof}
  This follows as a corollary of the previous theorem by induction in one direction and the pigeonhole principle in the other direction.

>>>>

\begin{document}

\begin{thm}
The Grassmann necklace $(I_1 \ldots I_n)$ from [previous theorem] is the Grassmann necklace for $M(W)$.
\end{thm}

\begin{proof}
We have shown that $(I_1,\dots,I_n)$ is a Grassmann necklace; it remains to check that this Grassmann necklace corresponds to the positroid $M(W)$.  Recall from the discussion following Definition~\ref{def:grassmann necklace}, and equation \eqref{basesofmatroids} in particular, that it suffices to show: 
\begin{itemize}
\item For each $i \in [n]$, $I_i$ is a basis for $M(W)$.
\item If $J$ is lexicographically smaller than $I_i$ with respect to $<_i$ (for any $i \in [n]$) then $J$ is not a basis for $M(W)$. 
\end{itemize}
The first point follows immediately: the algorithm is pairing each $j \in I_i$ with a unique propagator supported on that vertex, so $I_i$ is a basis for $M(W)$ by Corollary~\ref{lem basis as perm}.

Now fix $i \in [n]$, and write $I_i = \{a_1 <_i a_2 <_i \dots <_i a_k\}$. This imposes an ordering on the propagators $\cP$, where we label them $p_1, p_2, \dots, p_k$ such that $I_i(p_j) = a_j$ for each $j$. Define $\cP_m: = \{p_1,p_2,\dots,p_m\}$ for each $m\geq 0$, with $\cP_0 = \emptyset$.

We first note that if $a_{m-1}$ and $a_{m}$ are not consecutive vertices for some $m$, then for any intermediate vertex $a_{m-1}<_i x <_i a_{m}$ we must have $\Prop(x) \subseteq \cP_{m-1}$. This is because if any propagator supported on $x$ is still unassigned when the algorithm reaches $x$, we would have $x \in I_i$.

Let $J = \{j_1 <_i j_2<_i \dots <_i j_k\}$ be a $k$-set, and suppose that $J<_i I_i$ in the $<_i$-lexicographic order. Let $m$ be the minimal index such that $j_m <_i a_m$, i.e. the first entry at which $J$ and $I_i$ differ. We will show that $J$ must be a dependent set.

If $\Prop(j_m) = \emptyset$ then $J$ is automatically dependent since the corresponding column of $C(W)$ contains only zeros, so suppose henceforth that $\Prop(j_m) \neq \emptyset$. In particular, this implies that $m>1$, since if $m=1$ we would have $\Prop(j_1) \subseteq \cP_0 = \emptyset$.

We define a sequence of sets of propagators by
\[Q_0 = \Prop(j_m), \qquad Q_{l+1} = \Prop(I_i(Q_l))\quad \forall l \geq 0.\]
It is easily verified that this is an increasing chain of sets: indeed, if $q \in Q_l$ then $I_i(q) \in I_i(Q_l)$, and since $I_i(q)$ is one of the vertices that supports $q$ we must have $q \in \Prop(I_i(Q_l)) = Q_{l+1}$. The sequence therefore stabilizes in finitely many steps, since $\cP$ is finite.

We claim that $Q_l \subseteq \cP_{m-1}$ for all $l\geq 0$. The base case follows immediately from the observation above that $\Prop(j_m) \subseteq \cP_{m-1}$.

[induction step; see final page]

Let $Q:= \lim_{l\rightarrow \infty} Q_l$; by the above, this is well-defined and contained in $\cP_{m-1}$. We can now show that $J$ is a dependent set, by constructing a subset $U \subseteq J$ that satifies $|\Prop(U)| < |U|$.

Specifically, we take $U = I_i(Q) \cup \{j_m\}$. This is a subset of $J$ because $I_i$ and $J$ agree on their first $m-1$ terms and $Q \subseteq \cP_{m-1}$. Further, we have $j_m \not\in I_i(Q)$, so $|U| = |I_i(Q)| + 1 = |Q| +1$. On the other hand, since $\Prop(j_m) \subseteq Q$, we have
\[\Prop(U) = \Prop(I_i(Q)) \cup \Prop(j_m) = Q\]
by the construction of $Q$, and hence $|\Prop(U)| = |Q| < |U|$. Thus $J$ is a dependent set, and the Grassmann necklace $(I_1,\dots,I_n)$ constructed from $W$ does indeed define the positroid $M(W)$.
\end{proof}

\newpage

Half-written induction step:

Before proceeding with the induction, we make an observation about the configuration of propagators in $W$: given that $\Prop(j_m) \subseteq \cP_{m-1}$, we either have a propagator $r$ as in Cases 1 or 2 below, or we are in Case 3.

\vspace{3cm}

In Cases 1 and 2, let $r$ be the outermost of such propagators. In Case 3, we repeat the analysis with earlier vertices:

\vspace{3cm}

Eventually we either reach the starting vertex $i$ (in which case the sequence stabilizes at $Q_l = \cP_{m-1}$), or we reach some index $l_0$ such that $Q_{l_0} \subseteq \cP_{m-1}$ and $Q_{l_0}$ admits a propagator $r$.

Thus the result is true for all $l \leq l_0$; now suppose that $Q_l \subseteq \cP_{m-1}$ for some $l \geq l_0$ and consider $Q_{l+1} = \Prop(I_i(Q_l))$.

Then $q$ is supported on some vertex $a_t$, with $a_t = I_i(p_t)$ for some $p_t \in Q_l$. If $q$ lies to the left of $p_t$, then $q$ must have been assigned before $p_t$, i.e. $q \in P_{m-1}$.

If $q$ lies to the right of $p_t$,
